Simulating an Isolated Dwarf Galaxy with Multi-Channel Feedback and Chemical Yields from Individual Stars

### 2018-09-18 main.tex
final edits and reference updates

Using star-by-star modeling, we capture in more detail the stellar yields from individual stars released over their lifetime. This includes yields from massive and AGB stellar winds, and supernovae (SNe). In addition to better capturing how individual metal species enrich the ISM, this allows us to chemically tag individual stars. This ability opens an exciting new channel for testing models of galaxy evolution by leveraging current and ongoing observations probing the detailed distributions of chemical abundances of stars in the Milky Way and Local Group, such as APOGEE and APOGEE2 \citep{APOGEE2010,APOGEE}, the Gaia-ESO survey \citep{Gaia}, and GALAH \citep{GALAH}. This paper is the first in a series examining in detail the role that individual components of multi-channel stellar feedback play in galaxy dynamical and chemical evolution. We describe each mode of our multi-channel feedback 
in detail in Section~\ref{sec:methods}, describe their implementation in an isolated dwarf galaxy simulation in Section~\ref{sec:IC}, show results from this simulation in Section~\ref{sec:results}, discuss the results in Section~\ref{sec:discussion}, and conclude in Section~\ref{sec:conclusion}.

Readers who may want to only briefly skim (or skip) over the details of our included physics are advised to just read the beginning of Section~\ref{sec:methods}, which contains a complete---yet brief---summary of the included physics.

@@

At our adopted metallicity, nitrogen is predominantly produced in AGB star winds, with very little production in core collapse SNe and winds from more massive stars. Therefore, nitrogen is injected into the ISM with significantly less energy ($v \sim 10$~km~s$^{-1}$) than elements produced in SNe, like oxygen, ($v\sim 10^3$~km~s$^{-1}$). Given the variations in Fig.~\ref{fig:metal_slices}, the energetic differences between injection sources can drive abundance variations within the ISM of our dwarf galaxy. The regions most rich in N are sites of recent AGB winds that have yet to mix with the rest of the ISM. This suggests that metal mixing within the ISM (and also metal ejection from the ISM) is species dependent. A more detailed analysis is beyond the scope of this work, but we investigate this in detail in Emerick et. al. (in prep.).

@@

Leo P, the dwarf galaxy we approximate in our initial conditions, is extremely valuable as a gas-rich, star forming, low-mass dwarf galaxy, with an observable HII region, necessary for determining gas phase abundances, that is close enough to the Milky Way to conduct this experiment. Leo P retains $\sim$ 5\% $\pm$ 2\% of its metals, $\sim$ 1\% in stars and the rest in ISM gas \citep{McQuinn2015}. As discussed in Section~\ref{sec:chemical evolution}, more than 90\% of the tracked metals produced during our simulation no longer reside within the galaxy's disk, agreeing with observations. However, this is an evolving quantity that also depends on how much (if any) subsequent re-accretion of these metals occurs. Although more than half of these metals are expected to eventually re-accrete \citep{Christensen2016,Angles-Alcazar2017}, it is still possible that most of the metals that have been produced at these early times will remain outside the galaxy disk.

@@

Roughly half of massive stars live in binary pairs \citep{Sana2013}. Their interactions, primarily through mass transfer, can significantly alter their radiation properties and lifetimes. This can change both how much and how long these stars emit ionizing radiation, an important source of stellar feedback, and where and when these stars explode as SNe. This effect could be significant, but is rarely accounted for in galaxy evolution models, which are commonly based on calculations of single star evolution (e.g. STARBURST99). For example, \citet{Zapartas2017} finds that binarity extends the timescales over which core collapse SNe occur from a given star formation event, from a maximum time of $\sim$ 50~Myr to $\sim$~200~Myr. Although they find only $\sim 15\%$ of core collapse SNe explode after 50~Myr, this could still be an important effect. Properly accounting for the delay times due to variations in individual star lifetimes has already been shown to change the significance of feedback and influence galaxy metallicity properties \citep{Kimm2015}. Extending the lifetimes of these stars could increase the role of radiation feedback, however this may be less important as these additional photons are more likely to escape the galaxy \citep[e.g.][]{Ma2016-binary}.

@@

\item The ISRF of our galaxy varies strongly in both space and time by orders of magnitude. It is unclear how important these fluctuations are as a source of feedback, or if the affect can be approximated with a time-averaged radial profile. The importance of radiation feedback in our model will be investigated in a future work. (See Figs.~\ref{fig:ISRF} and \ref{fig:ISRF_2D}.)

@@

\item Finally, we present gas-phase oxygen and nitrogen distributions as examples to briefly demonstrate that there are marked differences in how individual metal species are distributed within the ISM of our galaxy. These variations could be tied to differences in elemental yields among different sources (for example, AGB winds vs. SNe), as suggested by \cite{KrumholzTing2018}. This will be explored in more detail in future work (Emerick et. al. \textit{in prep.}).
\end{enumerate}

>>>>

Using star-by-star modeling, we capture in more detail the stellar yields from individual stars released over their lifetime. This includes yields from massive and AGB stellar winds, and supernovae (SNe). In addition to better capturing how individual metal species enrich the ISM, this allows us to chemically tag individual stars. This ability opens an exciting new channel for testing models of galaxy evolution by leveraging current and ongoing observations probing the detailed distributions of chemical abundances of stars in the Milky Way and Local Group, such as APOGEE and APOGEE2 \citep{APOGEE2010,APOGEE}, the Gaia-ESO survey \citep{Gaia}, and GALAH \citep{GALAH}. This paper is the first in a series examining in detail the role that individual components of multi-channel stellar feedback play in galaxy dynamical and chemical evolution. \textbf{In \cite{Emerick2018b} we investigate the importance of ionizing radiation in regulating star formation and driving outflows in our galaxy. In \cite{Emerick2018c} we explore how individual metal species mix within the ISM and are ejected via galactic winds.}

We describe each mode of our multi-channel feedback 
in detail in Section~\ref{sec:methods}, describe their implementation in an isolated dwarf galaxy simulation in Section~\ref{sec:IC}, show results from this simulation in Section~\ref{sec:results}, discuss the results in Section~\ref{sec:discussion}, and conclude in Section~\ref{sec:conclusion}. Readers who may want to only briefly skim (or skip) over the details of our included physics are advised to just read the beginning of Section~\ref{sec:methods}, which contains a complete---yet brief---summary of the included physics.

@@

At our adopted metallicity, nitrogen is predominantly produced in AGB star winds, with very little production in core collapse SNe and winds from more massive stars. Therefore, nitrogen is injected into the ISM with significantly less energy ($v \sim 10$~km~s$^{-1}$) than elements produced in SNe, like oxygen, ($v\sim 10^3$~km~s$^{-1}$). Given the variations in Fig.~\ref{fig:metal_slices}, the energetic differences between injection sources can drive abundance variations within the ISM of our dwarf galaxy. The regions most rich in N are sites of recent AGB winds that have yet to mix with the rest of the ISM. This suggests that metal mixing within the ISM (and also metal ejection from the ISM) is species dependent. A more detailed analysis is beyond the scope of this work, but we investigate this in \textbf{detail in \cite{Emerick2018c}}.

@@

Leo P, the dwarf galaxy we approximate in our initial conditions, is extremely valuable as a gas-rich, star forming, low-mass dwarf galaxy, with an observable HII region, necessary for determining gas phase abundances, that is close enough to the Milky Way to conduct this experiment. Leo P retains $\sim$ 5\% $\pm$ 2\% of its metals, $\sim$ 1\% in stars and the rest in ISM gas \citep{McQuinn2015}. As discussed in Section~\ref{sec:chemical evolution}, more than 90\% of the tracked metals produced during our simulation no longer reside within the galaxy's disk, agreeing with observations. However, this is an evolving quantity that also depends on how much (if any) subsequent re-accretion of these metals occurs. Although more than half of these metals are expected to eventually re-accrete \citep{Christensen2016,Christensen2018,Angles-Alcazar2017}, it is still possible that most of the metals that have been produced at these early times will remain outside the galaxy disk.

@@

Roughly half of massive stars live in binary pairs \citep{Sana2013}. Their interactions, primarily through mass transfer, can significantly alter their radiation properties and lifetimes. This can change both how much and how long these stars emit ionizing radiation, an important source of stellar feedback, and where and when these stars explode as SNe. This effect could be significant, but is rarely accounted for in galaxy evolution models, which are commonly based on calculations of single star evolution (e.g. STARBURST99). For example, \citet{Zapartas2017} finds that binarity extends the timescales over which core collapse SNe occur from a given star formation event, from a maximum time of $\sim$ 50~Myr to $\sim$~200~Myr. Although they find only $\sim 15\%$ of core collapse SNe explode after 50~Myr, this could still be an important effect. Properly accounting for the delay times due to variations in individual star lifetimes has already been shown to change the significance of feedback and influence galaxy metallicity properties \citep{Kimm2015}. Extending the lifetimes of these stars combined with accounting for binary effects that change the luminosities of these stars \citep{Gotberg2017,Gotberg2018} could increase the importance of radiation feedback; however this may be less important as these additional photons are more likely to escape the galaxy \citep[e.g.][]{Ma2016-binary}.

@@

\item The ISRF of our galaxy varies strongly in both space and time by orders of magnitude. It is unclear how important these fluctuations are as a source of feedback, or if the affect can be approximated with a time-averaged radial profile. The importance of radiation feedback in our model is investigated in more \textbf{detail in \cite{Emerick2018b}.} (See Figs.~\ref{fig:ISRF} and \ref{fig:ISRF_2D}.)

@@

\item Finally, we present gas-phase oxygen and nitrogen distributions as examples to briefly demonstrate that there are marked differences in how individual metal species are distributed within the ISM of our galaxy. These variations could be tied to differences in elemental yields among different sources (for example, AGB winds vs. SNe), as suggested by \cite{KrumholzTing2018}. This is explored in more \textbf{detail in \citep{Emerick2018c}.}
\end{enumerate}
